---
title: "Comparing Female Lizard Lengths by Species"
author: "Amber McEldowney"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(here)
library(grid)
library(gridExtra)
library(effsize)
library(kableExtra)
```

This analysis compares total lengths for female lizards of two species CNUN (*Cnemidophorus uniparens*, the Desert Grassland lizard) and CNTE (*Cnemidophorus tessalatus*, the Colorado Checkered lizard) using Jornada Basin LTER lizards data.  The question addressed is: is there a significant difference in total length difference in mean total length between female lizards for the two species? The methods used are exploratory analysis, data visualization, data summary (finding mean, and standard deviation for total length, sample size, and each group), and the effect size will be found using Cohen's d test to determind the effect size between means for the two species.

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#read in data
lizards_length <- read_csv(here("data", "lizards.csv"))
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#Create and store a simplified version of the data frame that only includes variables spp, sex, and total_length
lizards_length %>% 
  select(spp,sex,total_length)

#Filter to only include observations for female lizards of species CNUN and CNTE
lizards_length_female <- lizards_length %>% 
  filter(sex == "F", spp == "CNUN" | spp =="CNTE")

#Create a finalized visualization comparing total lengths of female lizards for the two species.
lizards_length_female 
grid.labs <- c("Cnemidophorus uniparens", "Cnemidophorus tessalatus") 
names(grid.labs) <- c("CNTE", "CNUN")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Create a finalized visualization comparing total lengths of female lizards for the two species.
grid.labs <- c("Cnemidophorus uniparens", "Cnemidophorus tessalatus") 
names(grid.labs) <- c("CNTE", "CNUN")

ggplot(lizards_length_female, aes(x=total_length)) +
geom_histogram(aes(fill=spp),  show.legend = FALSE) +
  facet_wrap(~spp, labeller = labeller(spp=grid.labs)) +
  labs(x = "Length (mm)", y="Count of Lizards Observed", title = "Length Distributions of Female Lizards by Species")
```

**Figure 1.** This histogram compares total lengths of female lizards for the two species *Cnemidophorus uniparens*, the Desert Grassland lizard and *Cnemidophorus tessalatus*, the Colorado Checkered lizard using the Jornada Basin LTER lizards data.

```{r message=FALSE, warning=FALSE, include=FALSE}
# Faceted QQ plots: 
ggplot(data = lizards_length_female, aes(sample = total_length)) +
  geom_qq(aes(color = spp),
                 alpha = 0.5,
                 show.legend = TRUE
          ) 

```

**Table 1.** The following table includes the mean, standard deviation and sample size for female lizards CNUN (Cnemidophorus uniparens, the Desert Grassland lizard) and CNTE (Cnemidophorus tessalatus, the Colorado Checkered lizard).

```{r echo=FALSE, paged.print=TRUE}
#Prepare a finalized summary table that contains the mean and standard deviation for total length, and sample size, for each group (female CNUN and CNTE lizards)
cnun <- lizards_length_female %>% 
  filter(spp == "CNUN") %>% 
pull(total_length)

cnte <- lizards_length_female %>% 
  filter(spp == "CNTE") %>% 
  pull(total_length)

# Get the means & SDs for each: 
cnun_mean <- mean(cnun, na.rm = TRUE)
cnte_mean <- mean(cnte, na.rm = TRUE)
cnun_sd <- sd(cnun, na.rm = TRUE)
cnte_sd <- sd(cnte, na.rm = TRUE)

# t-test:
length_ttest <- t.test(cnun,cnte) 

# Get the effect size: 
length_effsize <- cohen.d(cnun,cnte, na.rm = TRUE)
# Make a table
length_summary <- lizards_length_female %>% 
  group_by(spp)  %>% 
  summarise("Mean Length (mm)" = mean(total_length, na.rm = TRUE),
                        "Standard Deviation" = sd(total_length, na.rm = TRUE),
            "Sample Size"= n())  

grid.labs <- c("Cnemidophorus uniparens", "Cnemidophorus tessalatus") 
names(grid.labs) <- c("CNTE", "CNUN")

length_summary %>% 

      kable(col.names = c("","Mean Length (mm)","Standard Deviation","Sample Size")) %>% 
kable_classic_2()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
length_clean <- broom::tidy(length_ttest)
```
There is a significant difference in length between Cnemidophorus uniparens and Cnemidophorus tessalatus, and a large effect size (Cohen's *d* = -2.41). t(`r round(length_clean$parameter,2)`) = `r round(length_clean$statistic,2)`, *p* = `r length_clean$p.value`.  There is a 95% chance that the actual mean for female *Cnemidophorus uniparens* lengths is between 113.1mm and 182.16mm and the actual mean for female *Cnemidophorus tessalatus* lengths is between 198.56mm and 292.22mm. 

**Citation** Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 
