---
title: "Exploring Environmentalist Identification and Political Affiliation Data"
author: "Amber McEldowney"
date: "12/7/2020"
output: html_document
---

In this analysis I will explore the question "is there an association between political party affiliation, and whether or not one self-identifies as and environmentalist"?  I will create a table with the data, and perform a chi-square test for independence to address this question, and discuss the results.
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(here)
library(dplyr)
library(janitor)
library(broom)
library(kableExtra)
```


```{r}
##Read in Conservation Survey data
cons_survey <- read.csv(here("data", "conservation_survey.csv"))
cons_survey_clean <- clean_names(cons_survey)
```


```{r}
##Data Wrangling
#Here I will keep only the variables needed for this task "ENV_ID" and "POL_OR"
pol_env <- cons_survey_clean %>% 
 select(env_id, pol_or)
```

```{r}
##Simplify Data Frame
#Add a new column that bins political response of either “conservative” (if POL_OR is 1 - 3) or “liberal” (if POL_OR 5 - 7). 

pol_env_name <- pol_env %>% 
  mutate(pol_name = case_when(pol_or %in% c("1","2","3") ~ "Conservative", pol_or %in% c("5","6","7") ~ "Liberal")) %>% 
  drop_na()
```

**Table 1.** This table shows the counts and proportions of responses to survey statement "I consider myself an environmentalist," (with outcomes Yes, No, and unsure) and illustrates the association between respondents’ political affiliation (binned “conservative” or “liberal” by self-identification) and if they consider themself an environmentalist.
```{r}
##Create a table
#Here I will create a finalized table that shows the counts and proportions of “liberal” and “conservative” survey respondents who responded “Yes”, “No” and “Unsure” to the statement “I consider myself an environmentalist.” 

pol_env_results <- pol_env_name %>% 
  janitor::tabyl(pol_name, env_id) %>% 
  rename("Yes" = "1", "No" = "2", "Unsure"="3") %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits=2) %>% 
  adorn_ns() %>% 
  group_by(pol_name)


pol_env_results %>% 
  kable(col.names = c("","Yes","No","Unsure")) %>% 
  kable_classic()



```

```{r message=FALSE, warning=FALSE, include=FALSE}
## Make contingency table and run chi-square
pol_env_counts<- pol_env_name %>% 
  janitor::tabyl(pol_name, env_id) %>% 
  rename("Yes" = "1", "No" = "2", "Unsure"="3") %>% 
  column_to_rownames(var="pol_name")
  
pol_env_counts %>% 
  kable(col.names = c("Yes","No","Unsure")) %>% 
  kable_classic()

pol_env_chi <- chisq.test(pol_env_counts)

pol_env_chi_tidy <- broom::tidy(pol_env_chi)
```

We retain the hypothesis that there is a significant association between political preference and identifying as an environmentalist - the two variables are not independent ($\chi$^2^(`r pol_env_chi_tidy$parameter`) = `r round(pol_env_chi_tidy$statistic,2)`, *p* = `r round(pol_env_chi_tidy$p.value, 18)`). This means, there is an association between political party, and self-identification as an environmentalist in this dataset.

**Data citation**: Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c

